### 2015-05-06 summary/ABPIC-PAPER/msv1.tex
finally starting to write paper

\documentclass[iop]{emulateapj}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\shorttitle{Short Title}
\shortauthors{Zhou et al.}

\begin{document}
\title{Cloud structure of planetary mass companion AB Pic-b}
\author{Yifan Zhou, Daniel Apai, ...}
\affil{University of Arizona}

\begin{abstract}
...
\end{abstract}

\keywords{kw1, kw2, ...}
\maketitle
\section{Introcduction}

\section{Observation}
\section{Data Reduction}
\section{Result}
\section{Discussion}
\acknowledgments{
  Acknowledgments. 
}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: