### 2015-08-14 summary/letter/letter_v1.tex
ok, roll subtraction

\documentclass[apj]{emulateapj}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\bibliographystyle{apj}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\shorttitle{Short Title}
\shortauthors{Zhou et al.}

%%% new command %%%
\newcommand{\ima}{\texttt{ima} files }
\newcommand{\flt}{\texttt{flt} files }
\newcommand{\eps}{$\mathrm{e}^{-}/\mathrm{s}$}
\begin{document}
\title{Variability of Planetary Mass Companion 2M1207 B}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider, ...}
\affil{University of Arizona}

\begin{abstract}
...
\end{abstract}

\keywords{kw1, kw2, ...}
\maketitle
%
\section{Introcduction}

\section{Observation}
We took high contrast direct imaging observation of 2M1207
system on UT 2014 April 11 using HST.
(\textit{HST} Proposal GO-13418, PI: D. Apai). The data were acquired
with the Wide Field Camera 3 (WFC3) in filter F125W
and F160W using the $256\times256$ sub-array mode. We observed
2M1207 system in continuous 6 HST orbits and obtained a data-set that
have a temporal resolution of $\sim1.5$ minutes and baseline of $\sim
8$ hours.

The observation was initially designed to apply standard roll
subtraction to remove the primary star. The telescope roll angles for
data taken in orbit 1, 3, 5 and those taken in orbit 2, 4, 6 have a
difference of $25^{\circ}$, thus the companion would appear at
different position relative to the primary 
and the primary in the two data set can be subtracted from each other.
In each orbit we took a sequence of 13 SPARS25 exposures with
NSAMP=10, alternating between F160W and F125W filters, with 2-3
identical exposures in one exposure sequence. To improve PSF
sampling and reduce the risk caused by bad pixels, we applied standard 4
point dithering. In each orbit, exposures were taken with dithering
position 1 -- 4 sequently. To sum, we obtained 70 multi-acuum images
for filter F125W and 64 images for filter F160W with exposure time of
88.4 seconds for both two filters.

\section{Data Reduction}

Through the whole analysis, we used the \flt that were
produced by WFC3's \texttt{calwfc3} pipeline. The \flt were reduced
with calibration steps including dark current correction, non-linearity
correction, and flat field correction. A further step of up-the-ramp
fit was applied to combine all non-destructive reads and remove cosmic
rays. Although \cite{Mandell2013} stated that WFC3 IR time series
extract from {\flt} have a rms 1.3 times larger than that obtained
from {\ima}, \flt keeps all pixels of the image of 2M1207 A
unsaturated, while in 80\% of non-destructive reads the cores of 2M1207
A are saturated. Thus \flt help better align the images for primary
subtraction, and separate the flux of 2M1207 B from that of 2M1207A.

The small angular separation of 2M1207 A and B (as shown in Figure
\ref{fig:1}) makes precise primary star subtraction and photometry
very difficult. On the under-sampled WFC3 IR detector (plate scale
$\sim 0.13''\mbox{pixel}^{-1}$ \cite{dressel2012wide}), the primary
and the secondary only separate by $\sim 6$ pixels, which is about 5
times of the FWHM of the PSF. In addition, under-sampling causes
significant artifacts when shifting the images for registration no
matter what interpolation method is used.

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{WFC3 F160W images of 2M1207A system. Upper: original image,
    lower: residual model and primary PSF subtracted. The image of
    2M1207 B is overshadowed by the halo of the image of 2M1207A and
    can be hardly seen from the original image. With a hybrid PSF
    (residual + Tiny Tim PSF) subtraction, the image of 2M1207 B is
    clearly displayed.}
  \label{fig:1}
\end{figure*}

\subsection{Tiny Tim PSF Photometry}
To over come the difficulty, we make use of the Tiny Tim PSF simulator
to pursue high precision photometry under this extreme
circumstance. Tiny Tim can produce model PSF based on the filter,
spectrum of target, focus status, and the telescope jitter. One
significant advantage of Tiny Tim PSF over empirical PSF is that Tiny Tim can
produce Nyquist sampled PSF, which makes image shifting and interpolation
rather straight forward. One essential disadvantage of Tiny Tim is
that the model PSFs have significant systematic
errors. \cite{Biretta2014} demonstrated that the diffraction
spikes and coma are not well simulated with Tiny Tim for WFC3 IR
images. However, using the 6-orbit long time series, we are able
to well characterize the difference of model PSFs and empirical PSFs.
By modeling the difference and combining it with model PSFs, we can
produce high quality hybrid PSFs to significantly improve the PSF
simulation and PSF photometry.

We started the reduction by making bad pixel masks for every
image. Pixels with data quality flags ``bad detector pixels'' (DQ = 4),
``unstable response'' (DQ = 32), and ``bad or uncertain flat value'' (DQ =
512) were masked out and excluded from further analysis as suggested
by previous transit exoplanet
spectroscopic observations\citep[e.g.][]{Berta2012, Kreidberg2014}.

We then produced a list of 10x over-sampled PSFs based on the
positions and spectrum \citep{Bonnefoy2014} of 2M1207 A with different
telescope jitter ranging from 0 to 50 $\mbox{mini-arcsec s}^{-1}$ in
both $x$ and $y$ directions. We used the new set of Tiny Tim
parameters provided by \cite{Biretta2014} to better model the cold
mask, diffraction spikes, and the coma. The focus parameters were
calculated using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}. The
PSFs were registered with the original image by searching on a dynamic
grid and minimizing the difference of observation and model PSF at a
region centered on 2M1207 A with the image of 2M1207 B mostly
excluded. The best fitted primary star position and telescope jitter
were obtained in this step. Then, we generated the PSF for 2M1207 B
based on its position on detector, spectrum \citep{Patience2010}, and
the telescope jitter that had been obtained above. In the final step,
we combined the two PSFs together. We fixed the position of the
primary as it had been fitted in the first step and fitted for the
amplitudes of the two PSFs and the precise position of the
secondary. Since the total fluxes of the model PSFs were normalized to
unity as default, the fluxes of 2M1207 A and B were solely represented
by the amplitude of the two PSFs coordinately.

Because of the systematic errors of the Tiny Tim PSF, the quality of the
fitting is not optimistic. The reduced $\chi^{2}$ values were at level
of $\sim10$. However, we found that the residual patterns were very
stable among the image that have the same dithering position and
telescope rolling. Therefore, we median combined the residual images
that have the same dithering position and telescope rolling angle to
construct 8 residual models (4 dithering position $\times$ 2 telescope
rolling angle) for each filter. We pre-subtract the corresponding
residual from the original image and repeated the procedure that
listed above. Using this extra-step, the reduced $\chi^{2}$
greatly decreased to around unity.

\subsection{Uncertainty analysis}

\subsubsection{White noise}
The difficulties of high contrast photometry is finally exhibited by
the error bars of the measurement. The scatterings of the photometric
measurements are dominated by photon noise.  We estimated the photon
noise by calculating the total flux within a 3-pixel (0.39 arcsec)
aperture centered on 2M1207 B. The photon noises for F125W and F160W photometry are
2.1\% and 1.2\%. These calculation results are at the same level of
the scattering of the photometric time series.

\subsubsection{Flat field uncertainties}
In our observation, 2M1207 B were taken exposure at 8 different spots
on the detector (2 rolls $\times$ 4 dithering
positions). The uncertainties in the flat field are potential sources of variation. The uncertainty of WFC3 IR pipeline flat field
is $\sim 1\%$ \citep{dressel2012wide} and our time resolved observation for another target
demonstrate that aperture photometry for exposures taken at different
dithering position have scatterings at the same level.

Flat field uncertainties have smaller effects on PSF photometry since
multiple pixels are fitted simultaneously. To verify this, we multiplied
an artificial flat field error mask that is a uniform distributed
Gaussian noise array with mean of 1 and sigma of 1\% to every image,
and repeat the PSF photometry routine with the error-applied
images. We obtained almost same photometry measurements as the
original.

However, PSF profile changes with different exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. Because of these factors, The
fluxes for both 2M1207 A and B are clearly correlated with the
dithering position. To reduce the correlation, we normalize each group of exposures
that have the same dithering position and roll angle individually --
we took the median of the fluxes that were measured from these
exposures as normalization factors and devided them from every photometric
measurement. Because the normalization factor for each group of
exposures is calculated across the whole observation, this
normalization step have negligible impacts on variability analysis.

\subsubsection{Validity tests for the variability}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{systematics}
  \caption{Normalized light curves for 2M1207B with filter F125W
    (left) and F160W (right). Original photometric measurements are
    plotted with gray crosses. The two halves of data are binned
    individually, and plotted with red points (first half) and blue
    squares (second half). The F160W light curves are plotted with two
  fitted sinusoidal curves, The green curve is fitted with
  the period fixed to be same as the F125W and the purple curve is
  fitted with all parameter set free.}
  \label{fig:2}
\end{figure*}

Because the aim of this letter is to search for temporal change, it is
essential to exclude the possibility that variability is systematic
artifacts.

 We fitted sinusoidal waves to both of the F125W and F160W light
 curves to preliminarily quantify the variability. The periods of the
 best fitted sine function of the two light curves are similar, 10.9
 hour for F125W and 9.3 hour for F160W, and the periods fitted here
 are not related with any WFC3 instrumental periods. The similarity of
 the period measurements is a proof that the variability is physical.

 Due to image persistence effect, the first orbit of HST observation
 is often problematic, and is neglected by several analyses
 (\textbf{citation?}). In our analysis, the fluxes measured from 2M1207 A are
 significantly lower in the first lower (Figure \ref{fig:3}) than
 those measured from rest fo the orbits. The exposure levels at the
 pixels of the peak of the PSF of 2M1207 B are less than 10000 e$^{-}$
 for both two colors, thus image persistence is less of an issue for
 2M1207 B. To further exclude this effect, we fitted sinusoidl waves
 to the light curves without using the first orbit, and obtained
 similar results.

 To rule out the possibility that the variability is introduced by
 data reduction, e.g., the normalization step, we split the data into
 two halves -- the first half is the data taken at dithering position 1
 and 3, and the second half is that for ditheirng position 2 and
 4. For each half, we repeat the analysis independently. For both of
 F125W and F160W, two halves demonstrated similar trend of
 variability as shown in Figure \ref{fig:2}.
 % \begin{itemize}
% \item the source of the scattering -- 
% \item flat field uncertainty -- adding artificial flat field error mask
% \item validity of the variability\\
%   -- both filter demonstrate similar period of variability\\
%   -- split the data into two half, the trend in two halves of data
%   looks similar.
%   -- fix the position does not change the light curve
% \item uncertainty of sinusoidal curve fit
%   -- carry out a mcmc fit? is it worth doing this.
% \end{itemize}

\section{Result}

\begin{figure*}
  \centering
  \plotone{sineCurveFit_binCombined}
  \caption{Normalized light curves for 2M1207 B (upper) and A (lower)
    with filter F125W (left) and F160W (right). Individual photometric
  measurement are plotted in gray crosses and binned photometry are
  plotted with red points. Best fitted sinusoidal waves are plotted
  with blue solid lines.}
  \label{fig:3}
\end{figure*}

We obtained high signal to noise photometry series for both 2M1207 A
and B (Figure \ref{fig:3}). On average, the photometric contrast is $6.52\pm0.03$ for
F125W and $5.77\pm0.02$ for F160W. The difference of F125W contrast
from that measured in J-band \citep{Mohanty2007} and
F160W contrast from that measured with NICMOS F160W 
\citep{Song2006} is due to the different throughput profiles of the
filters.

Although the photometric time series demonstrate relatively large
scattering, both F125W and F160W light curves demonstrate sinusoidal
shape with clear temporal variability. We fitted sine waves to the two
light curves using least square optimization, and determined the
periods and variation amplitudes. The best fitted periods for F125W
and F160W are 10.9 and 9.3 hour correspondingly. The amplitudes for
the normalized light curves are 1.4\% and 0.9\% for F125W and F160W
light curves.

\begin{figure*}
  \centering
  \plottwo{periodDistr}{amplitudeDistr}
  \caption{Distributions for periods (left) and amplitudes(right) for the light
    curve of F125W and F160W. The bin size for histograms of period is
  0.25 hour and for that of amplitude is 0.5\%. Histograms are
  normalized in the way that total area of the histogram equals to
  1. In the right panel, Gaussian profiles are fitted to the
  histograms and plotted in solid lines.}
  \label{fig:4}
\end{figure*}

To constrain the uncertainties of the least square optimization
results, we used a Monte Carlo(MC) method to improve the fitting. We
generated a series of random Gaussian noises with the standard deviation same as
the photon noise, added them to the original light curves, and applied
least square fit to the new light curves. We repeated above routine for 10000 times and obtained
the distribution of the fitting parameters. The distributions for the
periods and the amplitudes for F125W and F160W light curves are
shown in Figure \ref{fig:4}.

The distributions for the periods demonstrate long tail shaped towards
long period. The peaks of the two distributions separated by $\sim 1$
hour. The difference of the peaks are within $1-\sigma$. Combing the
measurements from the two light curves, we conclude the rotation period
of 2M1207B to be $9.0_{-1.5}^{+2.5}$ hours.

The variation amplitudes in the two bands have significant
difference. The distributions of the amplitudes are well described by
Gaussian profiles. To fit the histogram to a Gaussian function, we
determined that amplitude distribution of F125W peaks at 1.5\% with a
standard deviation of 0.3\%, and that of F160W have mean and standard
deviation of 0.9\% and 0.2\% correspondingly. The peaks of the two
histograms separated by more than $2-\sigma$. The variation amplitude
of F125W light curve is 1.67 times of that of F160W light curve.

\section{Discussion}
\begin{itemize}
\item a data-reduction pipeline is developed to obtain high precision
  photometry measurement from high contrast WFC3 IR data. For a
  contrast of $\sim 7$ magnitudes at an angular separation of
  $\sim0.7''$, we obtained photometry measurement for 2M1207 B at
  precision of about 2-3\%. 
\item time variability for 2M1207 B was discovered, the light curve of
  2M1207 B for both two colors can be fitted with a T=10.7 hr
  sinusoidal curve.
\item constrain on the amplitude of the amplitude of the
  variability. Large amplitude can be excluded. Constraints on the
  inhomogeniety of the cloud coverage can be inferred -- small
  thickness variance...
\item ? The atmosphere and cloud structure of 2M1207 B. How does it
  compare to the brown dwarf.
\end{itemize}
\bibliography{ref.bib}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: