### 2015-09-15 summary/letter/letter_vAD.tex
its been a long time since last commitment

\documentclass[apj]{emulateapj}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\bibliographystyle{apj}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\shortauthors{Zhou et al.}

%%% new command %%%
\newcommand{\ima}{\texttt{ima} files }
\newcommand{\flt}{\texttt{flt} files }
\newcommand{\eps}{$\mathrm{e}^{-}/\mathrm{s}$}
\newcommand{\tinytim}{\textit{Tiny Tim}}
\newcommand{\bpic}{$\beta$ Pic}
\newcommand{\vsini}{$v\sin i$}
\newcommand{\mjup}{M$_{\mbox{Jup}}$}
\begin{document}

\title{Discovery of Rotational Modulations in the Planetary-mass
  Companion 2M1207b: A Slow Rotation Period and Cloud Height in a Low
  Gravity Atmosphere}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider}
\affil{Steward Observatory, the University of Arizona}
\author{Mark Marley}
\affil{NASA Ames Research Center}

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, detection of periodic light curve variations
  can directly probe the rotational periods of ultra-cool objects.

  We present here, for the first time, time-resolved high-precision
  photometric measurements of a planetary-mass companion, 2MASS1207b,
  to a brown dwarf primary. Using HST/WFC3 and point spread
  function %in
  combination with two spacecraft roll angles we detect photometric
  modulations in the light curve. The amplitude is 0.9\% in the F160W
  and 1.5\% in the F125W filters; we find a consistent period and
  similar phase in both bands. Joint fit to the light curve in both
  bands suggest a period of $10.2^{+0.9}_{-0.8}$ h. The relative
  amplitudes in the two filters are very similar to that found in a
  recent study of a field (high-gravity) L-dwarf, suggesting that the
  cloud structures that introduce the photometric modulations are
  similar in high- and low-gravity objects. Importantly, our study
  also measures, for the first time, the rotational period for
  directly imaged planetary-mass companion.
\end{abstract}

\keywords{brown dwarfs -- planets and satellites: atmospheres -- planets
  and satellites: individual (2M1207b) -- techniques: photometric}
\maketitle
%
\section{Introduction}

% Cloud properties in high- and low-gravity objects -- thick clouds: an
% explanation for very red and faint near-infrared fluxes of directly
% imaged planets
Presence of condensate clouds is one the most unique and essential
features of the ultra-cool atmosphere of direct imaged exoplanet and brown
dwarfs. The formation and properties of condensate clouds of are
intensively studied in recent years \citep[e.g.][]{Ackerman2001,
  Helling2008, Allard2012}, and great improvement has been
achieved in understanding the cloud behavior across different spectral
types, especially in the explanation of L-T transition \citep{Marley2010}.
Surface gravity is suggested to be another key parameters 
in defining cloud structures. Low surface gravity objects (e.g. HR8799 bcd, 2M1207
b) have significantly redder color compared to their spectral type
matched brown dwarfs, which are best explained by the existence of 
thick clouds \citep{Barman2011, Skemer2011, Skemer2012}. However, due to lack of observational constraint, the
behavior of clouds in the atmospheres of low surface gravity objects
are not as well as modeled as high surface gravity counterpart.

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
\citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014,Metchev2015}. \cite{Kostov2013}
demonstrated the great potential of these techniques in the study of the cloud properties of
direct imaged exoplanets and planetary mass companions. However, high
brightness contrast makes it much more challenging for direct
imaged exoplanet and planetary mass companions than brown
dwarfs to acquire high-precision light curves.

2M1207 b \citep{Chauvin2005} is a planetary mass companion that has mass of
$\sim 4M_{\mathrm{Jup}}$ and an angular separation of $0.78''$ to its host
2M1207 A,
an M8 brown dwarf. Relative moderate contrast (6.7 mag in J band, \cite{Mohanty2007}) makes it one
the most feasible planetary mass companions for high-precision
time-resolved observation and the best starting points for developing
strategies for time-resolved observation and data reduction in high
contrast case.

In this {\em Letter} we present the first, high-cadence, high-precision,
time-resolved {\em Hubble Space Telescope} (HST) photometric time
series of 2M1207 b, a directly imaged planet or
planetary-mass object. We successfully detect rotational modulation
and measure the amplitudes in two bands and determine the rotational
period.

\section{Observation}

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{WFC3 F160W images for 2M1207 system. {\em Upper panel}:
    original image, the position of 2M1207b is indicated using white
    circle.   {\em Lower panel:} residual image -- after the
    subtraction of the hybrid PSF, 2M1207 B is detected at a high significant level.}
  \label{fig:1}
\end{figure*}

We obtained direct images of the 2M1207A+b
system on UT 2014 April 11from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3 \citep[WFC3,
][]{Kimble2008} in the frame of the HST Proposal GO-13418 (PI:
D. Apai). We acquired the observations in filters F125W
($\lambda_{\mbox{pivot}}$ = 1245.9 nm, full width at half maximum
(FWHM) = 301.5 nm) and F160W ($\lambda_{\mbox{pivot}}$ 1540.52, FWHM =
287.9 nm), roughly corresponding to the J and H bands. The WFC3 pixel
scale is $\approx$13mas. We used the $256\times256$ pixels sub-array
mode to avoid memory dumps during the observations.  In order to
provide a near-continuous coverage for detecting modulations we
observed the 2M1207 system in six consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58 minutes long Earth
occultations every 94 minutes.

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbits 1, 3, and 5, and those taken in
orbits 2, 4, and 6 differed by $25^{\circ}$. At the separation of
2M1207b this angle difference corresponds to a displacement of
$0.34''$ or 2.75 and 2.30 resolution elements in F125W and F160W,
respectively. In each orbit we took thirteen SPARS10 non-destructive
read-outs with NSAMP=10, alternating between F160W and F125W filters,
with 2--3 identical exposures in one exposure sequence. To improve
sampling and reduce the risk that point spread funciton (PSF) lays on
bad pixels, we applied a 4-point dither pattern with differential "X/Y"
offsets of 1.375" in the detector frame, providing optimal
non-integral (half pixel) stepping of 10.5 and 8.5 pixels in F125W and
F160W, respectively. Over the 6 orbits, we obtained 70 images with 10
non-destructive read-outs in F125W and 64 images in F160W with
exposure time of 88.4~s in each filters.

\section{Data Reduction}

%Although \cite{Mandell2013} stated that WFC3 IR time series
%extract from {\flt} have a rms 1.3 times larger than that obtained
%from {\ima}, \flt keeps all pixels of the image of 2M1207 A
%unsaturated, while in 80\% of non-destructive reads the cores of
%2M1207 A are saturated. 
%Thus \flt help better align the images for
%primary subtraction, and separate the flux of 2M1207 B from that of
%2M1207A.

%The small angular separation of 2M1207 A and B (as shown in Figure
%\ref{fig:1}) makes precise primary star subtraction and photometry
%very difficult. 
%On the under-sampled WFC3 IR detector (plate scale
  %$\sim 0.13''\mbox{pixel}^{-1}$ \cite{dressel2012wide}), the primary
%and the secondary only separate by $\sim 6$ pixels, which is about 5
%times of the FWHM of the PSF. In addition, under-sampling causes
%significant artifacts when shifting the images for registration no
%matter what interpolation method is used.
 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoidal curves are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of variation. In upper left panel, green line is sinusoidal
    wave fitted with all parameter set free, and purple line is
    sinusoidal wave fitted with period set the same as that of
    F125W. {\em Middle}: sinusoidal waves fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinusoidal curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}
\end{figure*}

\subsection{Photometry}

We started the reduction from the the \flt files produced by the
WFC3's \texttt{calwfc3} pipeline. We did not opt to use the more
advance \ima pipeline products because these provided less information
on 2M1207A, which saturated after the first few read-outs.  The \flt
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs also considering
cosmic rays. From the beginning, pixels with data quality flags ``bad
detector pixels'', ``unstable response'', and ``bad
or uncertain flat value'' were masked out and excluded from
further analysis as suggested by previous transit exoplanet
spectroscopic observations\citep[e.g.][]{Berta2012, Kreidberg2014}.

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifact structures are generated
by image interpolation and shifting. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist or better
sampled PSF, but systematic errors of \tinytim{} PSF for WFC3 limits
its ability in high precision photometry\citep{Biretta2014}. However,
we are able to fully characterize the difference of model and observed
PSFs with 6 orbits time-resolved observation data. To obtain robust \tinytim{} PSF
photometry, we designed a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to improve model of the cold mask,
OTA spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}. To
register \tinytim{} PSF to observed PSF of 2M1207A, we moved the
over-sampled PSF over a coordinate grid (gird size=0.001 pixel) using
cubic interpolation and minimize the rms difference of the
observed and re-binned \tinytim{} PSFs over a region centered on 2M1207A with a
5-pixel-radius aperture centered on 2M1207b excluded. Then we introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b simultaneously
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs were very stable for
given PSF positions. Therefore, at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
observationally derived correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and fitted the three components together. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Relative photometry is acquired from the scaling
parameters of the model PSFs.

PSF profiles change with exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. We found a correlation of
photometry with PSF positions on detector frame for both 2M1207 A and
b. Correction were made by normalizing each group of
exposures that have the same dithering position and roll angle
individually -- we took the median of the fluxes that were measured
from these exposures as normalization factors and divided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation,
this normalization step has negligible impacts on variability
analysis.

  \begin{figure*}
  \centering
  \plotone{sineCurveFit_binCombined}
  \caption{Normalized light curves for 2M1207 B (upper) and A (lower)
    with filter F125W (left) and F160W (right). Individual photometric
  measurement are plotted in gray crosses and binned photometry are
  plotted with red points. Best fitted sinusoidal waves are plotted
  with blue solid lines.}
  \label{fig:3}
\end{figure*}

\subsection{Uncertainty Analysis: White noise}

First we estimated the photon noises for the photometry of 2M1207b. We
propagated the photon noises of every single pixel that were
calculated from count rates and detector gain, to the PSF fitting
results. The photon noises for photometry in F125W and F160W are
1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b were fitted simultaneously in the
PSF photometry procedure, the uncertainties of photometry and position
for the primary and secondary are coupled. Imperfection of position
measurement of 2M1207A could potentially affect the photometry of
2M1207b. We used a Monte Carlo (MC) method to
evaluate the overall systematic of the PSF fitting. We applied the
PSF photometry to images that were added with random Poisson noises and
repeat the photometry procedure for 1000 times. From the distribution of the
result, the uncertainty for F125W and F160W photometry are 
1.34\% and 1.12\%, respectively. We conclude that the white noise of
our observation is dominated by photon noise.

\subsection{Uncertainty Analysis: Flat field uncertainties}

2M1207b were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction are potential sources of variation as they may
introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is typically $\sim 1\%$
\citep{dressel2012wide}.

In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect a lower than 1\% uncertainty from the flat field
errors. To verify this, we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical light curve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}).

\section{Verification of Photometric Variations and Amplitude
  Estimate}

\begin{figure*}
  \centering
  \includegraphics[width=0.45\textwidth]{periodDistr}
  \includegraphics[width=0.45\textwidth]{amplitudeDistr}
  \caption{Distributions for periods (left) and amplitudes(right) for the light
    curve of F125W and F160W. The bin size for histograms of period is
  0.25 hour and for that of amplitude is 0.5\%. Histograms are
  normalized in the way that total area of the histogram equals to
  1. In the right panel, Gaussian profiles are fitted to the
  histograms of periods and plotted in solid lines.}
  \label{fig:4}
\end{figure*}

\subsection{Tests and Verification}

The light curves that resulted from our photometry show apparently
sinusoidal modulations, discussed in more details in
\S,\ref{Results}. To verify that these modulations are intrinsic to
the object and not result of our data reduction procedure or due to
instrumental changes, we carried out three different tests.

First, we fitted sine waves independently to the two filters to verify
the similarity of the signal in the two bands (Figure \ref{fig:2}). Inconsistent periods or
light curve shapes would argue against a genuine signal.  We found
that the periods of the best fit sine waves are similar,
$10.5^{+1.2}_{-1.3}$h for F125W and $9.1^{+1.1}_{-1.0}$h for
F160W. These periods are roughly consistent within the
uncertainty. Furthermore, these periods are not close to any
timescales over which HST or WFC3 are known to changes.

 As a second test we repeated the analysis neglecting the first
 orbit. The motivation behind this test was that, due to spacecraft
 thermal settling, the first orbit of HST observations is often
 slightly unstable and is often neglected in high-precision studies
 (\textbf{citation?}). Indeed, in our analysis 2M1207 A is
 significantly fainter in the first orbit (Figure \ref{fig:3}) than in
 the subsequent ones.
 Our analysis based on orbits 2--6 only found essentially identical
 results to our analysis of orbits 1--6, based on which we conclude
 that the first less reliable orbit does not affect our results
 significantly (Figure \ref{fig:2}, middle panel).

As a third test we explored whether a subset of images, perhaps
imperfectly normalized or correlated with specific instrument states,
could be driving the light curves into an apparently sinusoidal
shape. To test this possibility we split the data into two temporally
overlapping halves: sub-dataset one from dithering positions 1 and 3,
and sub-dataset two from dithering positions 2 and 4. For both
datasets we repeated our analysis independently.  For both of F125W
and F160W, two halves demonstrated similar trend of variability as
shown in Figure \ref{fig:2}.  Our analysis detected sinusoidal
modulations in {\em both} sub-datasets and in {\em both} filters, with
periods and amplitudes consistent with those derived from the complete
data set (\ref{fig:2}, upper panel).

 These tests demonstrate that the modulation seen in our data are
 consistently present in the different filters, in the different time
 segments of the data, and in data obtained in different dithering
 positions.

\subsection{Amplitude and Period Measurements}

To constrain the uncertainties of the least square optimization
results, we used a MC method to improve the fitting. We
generated a series of random Gaussian noises with the standard deviation same as
the photon noise, added them to the original light curves, and applied
least square fit to the new light curves. We repeated above routine for 100,000 times and obtained
the distribution of the fitting parameters. The distributions for the
periods and the amplitudes for F125W and F160W light curves are
shown in Figure \ref{fig:4}.

\section{Result}

\label{Results}
We present the first  high-resolution, high-cadence, and
high-precision photometry of a directly imaged planet or
planetary-mass companion. Our observations reveal a modulation in the
light curve of the $\sim 4 \mathrm{M_{{Jup}}}$ companion 2M1207b, the first detection
of modulations in directly imaged planetary mass objects.  The best fit
periods for F125W and F160W are 10.5 and 9.1 hour correspondingly. The
amplitudes for the normalized light curves are 1.45\% and 0.92\% for
F125W and F160W light curves.

We obtained high signal to noise photometry series for both 2M1207 A
and B (Figure \ref{fig:3}). On average, the photometric contrast is
$6.52\pm0.01$ mag for
F125W and $5.77\pm0.01$ mag for F160W. The difference of F125W contrast
from that measured in J-band \citep{Mohanty2007} and
F160W contrast from that measured with NICMOS F160W 
\citep{Song2006} is due to the different throughput profiles of the
filters.

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With 64\% confidence,
we estimated the 1-$\sigma$ range for the periods of F125W and F160W
to be $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ h,
respectively. The period of best fitted sine wave of F125W light
curve is 1.5h longer than that of F160W, and 0.3--0.4 h larger
1-$\sigma$ standard deviation. We also jointly fit the two band light
curves together forcing the periods of two sinusoids to be the
same. We derived a modulation period of $10.2^{+0.9}_{-0.8}$ h.

We discover that the variation amplitudes in the two bands are significantly
different. By fitting Gaussians to the MC fit result distributions, we
determined that variation amplitude of F125W is 1.45\% with a
standard deviation of 0.22\%, and that of F160W is 0.92\% with a
standard deviation of 0.20\%. The peaks of the two
histograms separate by more than 2-$\sigma$. The variation amplitude
for F125W is 1.58 times of that of F160W light curve.

\section{Discussion}

\begin{figure*}
  \centering
  \plottwo{rotationDiagram}{JH}
  \caption{comparison of 2M1207's rotation period and color change with brown
    dwarfs, \bpic{} b, and solar system planets. {\em Left}: period vs. mass
  plot for 2M1207b (red square), solar system planets and \bpic{} b
  (blue squares), and brown dwarfs (black circles, gray shade). The mass of brown
  dwarfs are assumed to be 30 \mjup{}. The gray rectangle that has
a  $\pm$15 \mjup range in $x$, and a $\pm \sigma$ of brown dwarf
periods range in $y$, indicates a
region where brown dwarfs most likely to appear in this
diagram. Rotation period monotonically decreases with the increase of
mass. {\em Right}: ratio
of variation amplitude in J and H band vs. spectral type for 2M1207b
and brown dwarfs. The colors of the points represent J$-$H magnitude,
and the sizes of the points are proportional to the J band variation
amplitudes. The gray dashed line is the result of a linear fit to
these points. Tight linear correlation of J- and H-band variation ratio
and spectral type is demonstrated in this diagram.}
x \label{fig:5}
\end{figure*}

% A fundamental result of our study is the direct determination of the
% rotation period of a directly imaged planetary-mass object. We
% convert the rotation period to equatorial velocity  by adopting a radius of 1 -- 1.4
% $R_{\mathrm{Jup}}$ for 2M1207b, and 1 $R_{\mathrm{Jup}}$ for field
% brown dwarfs with well defined rotation period from the study of
% \cite{Metchev2015}, and compare their rotation velocities with solar
% system planets and \bpic{} b in the left
% panel of Figure \ref{fig:5}.  The study by
% \citep[][]{Snellen2014} succeeded in measuring \vsini{} for \bpic{} b
% and demonstrated that it fits a trend defined by Solar System planets
% in which more massive planets have faster rotation rates. They suggested
% that this relation is linked to the accretion processes during planet
% formation.

A fundamental result of our study is the direct determination of the
rotation period of a directly imaged planetary mass object. In the
left panel of Figure~\ref{fig:5} we compare the rotation period of
2M1207b to the solar system planets, \bpic b the only other directly
imaged planet with an estimated period and measured {\em v sin i}, and
field brown dwarfs from the study of \citep[][]{Metchev2015}.  The
study by \citep[][]{Snellen2014} succeeded in measuring {\em v sin i}
for \bpic b and demonstrated that it fits a trend defined by Solar
System planets in which more massive planets have faster rotation
rates. The interesting finding that \bpic b, an exoplanet that formed
in a protoplanetary disk, follows this trend suggests a possibly
connection between planet mass, initial angular momentum, and
formation in a disk.

Excitingly, our measurement of the rotation period of 2M1207b, a
planet mass companion that has similar age to \bpic b, has a rotation
period that fits in the same trend, as well as majority of brown
dwarfs.  We note that 2M1207 b is most likely formed in the same way as
brown dwarfs by gravitational fragmentation. The result that objects
formed in different scenarios share the same trend of period vs. mass suggests that
rotation periods -- in absence of well-determined ages -- are not good tracers of the formation pathways and
may not contribute important evidence for a formation in a disk vs. in
a cloud core environment.
{\em Add younger BDs and this may change}
% Excitingly, our measurement of the rotation period demonstrates that
% 2M1207b, a planetary mass companion  with similar age to \bpic{} b, has a rotation velocity
% that fits in the same trend, as well as majority of brown
% dwarfs. We note that 2M1207b is most likely formed in the same way as brown
% dwarfs by gravitational fragmentation, which suggests
% that rotation periods are not good tracers of the formation pathways
% and may not contribute important evidence for a formation in a disk
% vs. in a cloud core environment.

Furthermore, our observations allow us to compare the relative
amplitudes in the J- and H-bands with the handful of brown dwarfs
for which high-quality near-infrared time-resolved observations have
been obtained. In the right panel of Figure \ref{fig:5}, we compare the
relative amplitude of J- and H-bands of 2M1207b and brown dwarfs
\citep{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014} that
have different spectral types and J$-$H colors.

We find a tight correlation between the spectral type of the object
and the J- to H-band variation amplitude ratio. In the right panel of
Figure \ref{fig:5}, we demonstrate that earlier spectral type
objects, independent of their surface gravity, have larger variations
at shorter wavelength than at longer wavelengths. Importantly,
although the J$-$H color of 2M1207b is significantly redder, its
relative variation amplitude ratio is almost identical to the matching
spectral type mid-L dwarf 2M1821.

This trend can be qualitatively explained using the toy model proposed
in \cite{Yang2014} within the spectral type range of mid-L to L-T
transition. Assuming that intensity modulations are introduced by a
refractory cloud whose upper layer locates at altitude $z$, the
variation amplitude ratio for J and H band can be expressed as
$\mathrm{Amp_{J}}/\mathrm{Amp_{H}} =
\exp\left[-(\tau_{\mathrm{J}}(z)/\tau_{{\mathrm{H}}}(z)\right]$,
where $\tau_{\mathrm{J}}(z)$ and $\tau_{\mathrm{H}}(z)$ are the
optical depths of upper cloud layer for J and H bands. The opacity for
J band is smaller than that for H, therefore from mid-L to L-T
transition, as the cloud upper boundary appears higher in the
atmosphere, the optical depth difference for the two bands gets
smaller, and the amplitude ratio decreases.  We note, that the late T
dwarf 2M2228 falls exactly on this trend, but whose variations are
likely to emerge from a different mechanism as their silicate clouds
are below the visible photosphere \citep[e.g.][]{Stephens2009}.

We infer that the upper cloud layer of 2M1207b is at a similar optical
depth as that of a matching spectral type field brown dwarf from the fact
that they share similar J- and H-bands light curve amplitude ratio.

%The
%two processes that atmosphere opacity at certain pressure level
%enhances and the cloud thickness increases with the decrease of surface
%gravity have to counterbalance with each other to maintain the optical
%depth of the cloud upper layer.

\section{Conclusions}
In summary, from our J- and H-band high precision, high-cadence
light curves we discovered sinusoidal modulations in the planetary mass
object 2M1207b. This is the first detection of rotational modulations
in a directly imaged planetary mass object.  The period is
$10.2^{+0.9}_{-0.8}$,  similar to that derived from {\em v sin i}
measurements for the direct imaged exoplanet \bpic{} b and
significantly longer than most field brown dwarfs with known rotation
periods. The modulation amplitude ratio of J and H band is almost
identical to one matching spectral type L5 dwarf, although they have
very different J$-$H colors.

Finally, we note that the observations presented here open an exciting
new window on directly imaged exoplanets and planetary-mass
companions. Our study demonstrates a successful application of high-cadence,
high-precision, high-contrast photometry with planetary mass
companion. We also show that these observations can be carried out simultaneously
at multiple wavelengths, allowing us to prove multiple pressure
levels. With observation of a larger sample and at multiple wavelengths, we will be
able to explore the detailed structures of atmospheres of directly
imaged exoplanets, and identify the key parameters that determine these.

\bibliography{ref.bib}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: